---
title: "Image Analyzes vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Image Analyzes vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(AnalyzeR)
library(magick)
library(tiff)
```

# Introduction

This vignette demonstrates how to use the image analysis package to process and analyze images. The package provides functions for thresholding, grouping, and batch processing of images. \# Loading the Package First, load the package:

```{r}
library(AnalyzeR) 
library(magick) # For loading test images
```

# Example Images

For this demonstration, we'll use the cerebellum images included with the package:

```{r}

# In your vignette
library(AnalyzeR)
library(magick)

#  direct URLs to the raw GitHub files
img1 <- image_read("https://raw.githubusercontent.com/kathrynnrhodes/CollaborativeProject/main/F_FXR_3_cerebellum_40x_1.tif")
img2 <- image_read("https://raw.githubusercontent.com/kathrynnrhodes/CollaborativeProject/main/F_FXR_3_cerebellum_40x_2.tif")
img3 <- image_read("https://raw.githubusercontent.com/kathrynnrhodes/CollaborativeProject/main/F_FXR_3_cerebellum_40x_3.tif")

#  proceed with your analysis
analysis_result <- analyze_image(img1, threshold = "60%")
# View the analysis results
print(analysis_result$formatted)
print(paste("Image dimensions:", analysis_result$width, "x", analysis_result$height))
# Display the processed image
analysis_result$processed_image


```

# Basic Image Analysis

let's analyze a single image with our 'analyze_image' function:

# Batch Processing

Now, let's demonstrate how to process multiple cerebellum images using the package's batch capabilities:

```{r}
# Group the cerebellum images
image_list <- list(img1, img2, img3)
grouped_images <- group_loaded_images(image_list, process_loaded_images)

# Display information about the grouped images
print(paste("Number of images in group:", grouped_images$count))

# Print basic info about each image
for (i in 1:length(image_list)) {
  info <- magick::image_info(image_list[[i]])
  cat(sprintf("Image %d: %d x %d pixels\n", i, info$width, info$height))
}
```

# Batch Analysis

Now we can analyze all the images in the group at once:

```{r}
# Analyze all images with the same threshold
batch_results <- batch_analyze_images(grouped_images, threshold = "65%")

# View the summary results
cat(batch_results$summary)

# You can also access individual results
print(batch_results$individual_results[[1]]$formatted)
print(batch_results$individual_results[[2]]$formatted)
print(batch_results$individual_results[[3]]$formatted)

```

# Visualizing Results

Let's create a simple plot to visualize the analysis results:

```{r}
# Extract percentage values for plotting
percentages <- sapply(batch_results$individual_results, function(x) x$percentage)
image_names <- sapply(batch_results$individual_results, function(x) x$image_name)

# Create a bar plot
barplot(percentages, names.arg = image_names, 
        main = "Percentage Differences by Image",
        xlab = "Image", ylab = "Percentage (%)",
        col = "skyblue")
abline(h = batch_results$average_percentage, col = "red", lwd = 2)
legend("topright", legend = c("Individual Images", "Average"),
       fill = c("skyblue", "red"))


```

# Advanced Usage

```{r}
# Example of customizing thresholds for different image types
light_image_analysis <- analyze_image(img1, threshold = "40%")
dark_image_analysis <- analyze_image(img2, threshold = "80%")

# Compare results
cat("Light image threshold (40%):", light_image_analysis$formatted, "\n")
cat("Dark image threshold (80%):", dark_image_analysis$formatted, "\n")

```

# Conclusion

This vignette has demonstrated how to use the main functions of the image analysis package for analyzing both individual images and batches of images. By adjusting thresholds and processing parameters, you can customize the analysis for different types of images and research needs.
